Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-08-29 Manuscript/manuscript.Rmd
finish methods

Random forest classification models were built to compare model performances between combinations of data resolutions, predictor types, and re-sampling techniques. To reduce bias due to random dataset generation, 10-fold cross validation was used to tune the number of annotations, *p*, to consider at each node. The number of trees that were aggregated for each random forest model was set to the default, n=500. All random forest algorithms were implemented in R using the *caret* package version 6.0 [@kuhn2012caret]. The performance of each model was compared using Matthews Correlation Coefficient (MCC), defined as <!--The MCC is, in essence, the correlation coefficient between the observed and predicted binary classifications [@brodersen2010balanced]. A value of 1 represents a perfect prediction, while a value of -1 represents total disagreement between prediction and observation. A value of 0 indicates no better than random prediction. The MCC is defined as-->

@@

where TP refers to the number of bins correctly identified as containing a TAD boundary (true positives), FP refers to the number of bins incorrectly identified as containing a TAD boundary (false positives), TN refers to the number of bins correctly identified as not containing a TAD boundary (true negatives), and FN refers to the number of bins incorrectly identified as not containing a TAD boundary (false negatives). Each of these quantities are obtained from the confusion matrix created by validating the model on the testing set. <!--Additional performance metrics including AUROC, accuracy, sensitivity, and specificity were [??? are?] reported in the supplementary materials. [??? Where is it?]-->

@@

Many genomic annotations, notably architectural proteins, tend to exhibit an extensive pattern of colocalization (correlation) [@van2014insulator]. To suitably reduce the feature space and prevent overfitting, we implemented recursive feature elimination to estimate the optimal number of necessary features to include in the model. The best performing subset was identified, ranging from 2 to the maximum number of features by powers of 2. We then aggregated the predictive importance of the union of the optimal set of features across chromosomes using the mean decrease in node impurity to determine the most common and top ranked annotations for predicting TAD boundaries.

@@

By leveraging the enrichment of top ranked annotations in our model, we established *predicted TAD boundary regions* as a collection of adjacent base pairs that were given a probability of 1 for being a potentially predicted TAD boundary. To guard against inflating the number of predicted TAD boundaries, we clustered the distances between the midpoints of the *predicted TAD boundary regions*. We then performed partitioning around mediods (PAM) to determine the *predicted TAD boundary region* that was most representative of the cluster. The midpoint, in this case a particular basepair, of the *predicted TAD boundary region* identified as the medoid was defined as a final *predicted TAD boundary*. This process was repeated throughout the chromosome-specific sections of the linear genome.

>>>>

Random forest classification models (the `caret` v.6.0 R package [@kuhn2012caret]) were built to compare model performances between combinations of data resolutions, predictor types, and re-sampling techniques. Ten-fold cross validation was used to tune the number of annotations, *p*, to consider at each node. [??? Isn't it used for preventing overfitting? TBD] The number of trees that were aggregated for each random forest model was set to $n=500$.

The performance of each model was compared using Matthews Correlation Coefficient (MCC), defined as <!--The MCC is, in essence, the correlation coefficient between the observed and predicted binary classifications [@brodersen2010balanced]. A value of 1 represents a perfect prediction, while a value of -1 represents total disagreement between prediction and observation. A value of 0 indicates no better than random prediction. The MCC is defined as-->

@@

where TP refers to the number of bins correctly identified as containing a TAD boundary (true positives), FP refers to the number of bins incorrectly identified as containing a TAD boundary (false positives), TN refers to the number of bins correctly identified as not containing a TAD boundary (true negatives), and FN refers to the number of bins incorrectly identified as not containing a TAD boundary (false negatives). Each of these quantities are obtained from the confusion matrix created by validating the model on the testing set. <!--Additional performance metrics including AUROC, accuracy, sensitivity, and specificity were [??? are?] reported in the supplementary materials.-->

@@

Many genomic annotations, notably architectural proteins, tend to exhibit an extensive pattern of colocalization (correlation) [@van2014insulator]. To suitably reduce the feature space and prevent overfitting, we implemented recursive feature elimination to estimate the optimal number of necessary features to include in the model. The best performing subset was identified, ranging from 2 to the maximum number of features by powers of 2. We then aggregated the predictive importance of the union of the optimal set of features across chromosomes using the mean decrease in node impurity to determine the most common and top ranked annotations for predicting TAD boundaries. [??? TBD, at which stage it was done]

@@

Using optimally trained models, we established *predicted TAD boundary regions* as a collection of adjacent bases predicted as TAD boundary with the probability of 1. [??? Do not understand from here:] To guard against inflating the number of predicted TAD boundaries, we clustered the distances between the midpoints of the *predicted TAD boundary regions*. We then performed partitioning around mediods (PAM) to determine the *predicted TAD boundary region* that was most representative of the cluster. The midpoint, in this case a particular basepair, of the *predicted TAD boundary region* identified as the medoid was defined as a final *predicted TAD boundary*. [??? To here. REWRITE!]
